---
title: "Assignment 1"
author: "Madeleine Levin, Jaida Nabayan, Michael Zajakowski Uhll"
date: "10/21/2021"
output:
   html_document:
     theme: readable
     toc: true
     toc_depth: 3
     toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Research question

What is the relationship between the presence of open space and other amenities in NYC boroughs and the number of people arrested during protests against police brutality in summer 2020? 

# Prior research



# Data

The sample population for this study is the entire citizen population of New York City in the summer of 2020. The analysis includes the following variables:

*NYC Arrests for Summer 2020
*Location of Public Spaces in New York
*NYC Borough Boundaries
*NYPD Sector Boundaries
*Libraries

## Load data

```{r, message=FALSE}
library(tidyverse)
library(tidycensus)
library(readxl)
library(knitr)
```

Bringing in New York Data
```{r, message=FALSE, results='hide'}
nyc_arrests <-read.csv(file = "data/NYC_Arrests_Summer2020.csv")
wifiparks <-read.csv(file ="data/WiFiPublicSpace.csv")
Boroughs <- read.csv(file ="data/NYCBoroughBoundaries.csv")
NYPDSectors <-read.csv(file = "data/NYPD_Sectors.csv")
```

Bringing in Census Data
```{r}
census2020 <-get_decennial(geography = "county",
                        year= 2020,
                        variables =c(pop= "P1_001N",
                                     race_pop = "P1_002N"),
                        output = "wide",
                        geometry = TRUE)
```




